---
title: "CTMC"
author: "Jair Andrade"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(doParallel)
library(doRNG)
library(dplyr)
library(ggplot2)
library(pomp)
library(purrr)
library(readsdr)
library(stringr)
library(tictoc)

source("./R_scripts/local_search.R")
source("./R_scripts/loglik_funs.R")
source("./R_scripts/plots.R")

folder <- "./saved_objects"
```

```{r}
N_val <- 5234
```

```{r}
dat_df <- Maryland |> 
  select(Cumberland) |> 
  mutate(time = as.numeric(row_number()), .before = everything()) |> 
  rename(y = Cumberland) |> 
  as.data.frame()
```

```{r}
rinit <- Csnippet("
  S = nearbyint(N - I_0 - R_0);
  E = 0;
  I = nearbyint(I_0);
  R = nearbyint(R_0);
  C = 0;") 

dmeas <- "
  lik = dnbinom_mu(y, 1 / par_inv_phi, C, give_log);
"

SEIR_step <- "
  double rate[3];
  double dN[3];
  rate[0] = par_beta * I / N; // transmission
  rate[1] = par_sigma;        // symptom onset
  rate[2] = par_gamma;        // recovery
  reulermultinom(1, S, &rate[0], dt, &dN[0]);
  reulermultinom(1, E, &rate[1], dt, &dN[1]);
  reulermultinom(1, I, &rate[2], dt, &dN[2]);
  S += - dN[0];
  E += dN[0] - dN[1];
  I += dN[1] - dN[2];
  R += dN[2];
  C += par_rho * dN[1];
"

dt <- 1/64

params <- c(par_beta = 1, par_sigma = 0.5, par_gamma = 0.5,
                            par_rho = 0.75, I_0 = 1, R_0 = 0.3 * N_val,
                            N = N_val, par_inv_phi = 1/15)

fixed_params <- params[names(params) %in% c("par_sigma", "par_gamma", "R_0",
                                            "N")]

pomp_mdl <- pomp(data = dat_df,
                 times = "time", 
                 t0 = 0,
                 rinit = rinit,
                 rprocess = pomp::euler(Csnippet(SEIR_step), delta.t = dt),
                 dmeasure = Csnippet(dmeas),
                 statenames = c("S", "E", "I", "R", "C"),
                 paramnames = c("par_beta", "par_sigma", "par_gamma", 
                                "par_rho", "N", "I_0", "R_0",
                                "par_inv_phi"),
                 accumvars = "C",
                 params = params,
                 partrans = parameter_trans(
                   log   = c("par_beta", "par_inv_phi", "I_0"),
                   logit = "par_rho"),
                 cdir     = ".",
                 cfile    = "CTMC_SEIR")
```

```{r}
pf <- replicate(10, pfilter(pomp_mdl, Np = 5000))
ll <- sapply(pf,logLik)
logmeanexp(ll, se=TRUE)
```



```{r}
ptb <- rw_sd(I_0 = ivp(0.02), par_beta = 0.02, par_rho = 0.02, 
             par_inv_phi = 0.02)

seed <- 94638
fn   <- "./saved_objects/local_search.rds"

ls_obj <- local_search(pomp_mdl, ptb, fn, seed, n_cores = 7)
```

```{r}
mifs_local <- ls_obj$result

mifs_local |> 
  traces() |> 
  melt() -> traces_df 

mif_traces(traces_df, c("par_beta", "par_rho", "I0", "par_inv_phi"), 
           "#344D77")
```

### Likelihood estimates

\hfill


```{r}
fn <- file.path(folder, "local_search_ll.rds")

ll_local_search_obj <- mif_ll(mifs_local, seed = 359084918, 
                              n_cores = 7, filename = fn, Np = 1e4)
```

```{r}
df <- bind_rows(ll_local_search_obj$ll_results)
```

## Global search

### Likelihood maximisation

\hfill

```{r}
mf1          <- ls_obj$result[[1]] 

set.seed(583088)

runif_design(
  lower = c(par_beta = 0.1, par_rho = 0, I_0 = 0.5,  par_inv_phi = 0.01),
  upper = c(par_beta = 3.0, par_rho = 1, I_0 = 30, par_inv_phi = 1.00),
  nseq  = 300) -> guesses

fn     <- file.path(folder, "Global_search.rds")
seed   <- 48341889

gs_obj <- global_search(guesses, fixed_params, mf1, fn, seed, 7)
```
```{r, message = FALSE, fig.height = 3.5}
mifs_global <- extract_mif_results(gs_obj)

mifs_global |>  
  traces() |> 
  melt() -> traces_df 

mif_traces(traces_df, c("par_beta", "par_rho", "I0", "par_inv_phi"), 
           "#344D77")
```

### Likelihood estimates


```{r}
fn          <- file.path(folder, "Global_search_ll.rds" )
ll_obj      <- mif_ll(mifs_global, Np = 1e4, 146417047, n_cores = 7,
                      fn)
```

```{r, message = FALSE, fig.height = 6.5, fig.width = 6.5, fig.align = 'center'}
loglik_df <- extract_ll_df(ll_obj)

loglik_df |> 
  filter(loglik > max(loglik)- 20, loglik.se < 2) |>
  bind_rows(guesses) |>
  mutate(type=if_else(is.na(loglik),"guess","result")) |>
  arrange(type) -> guess_result_df

pairs(~loglik + par_beta + I_0 + par_rho + par_inv_phi,
      data = guess_result_df,
      col=ifelse(guess_result_df$type == "guess", grey(0.8), "#344D77"),
      pch = 16, cex = 0.5)
```

### $\beta$ - Effective contact rate

```{r}
#===============================================================================
var_name <- "par_beta"
#===============================================================================
```


#### Initial likelihood surface

\hfill

\hfill

```{r, fig.height = 2.5, fig.align ='center'}
GBM_colour <- "#344D77"
raw_likelihood(guess_result_df, var_name, GBM_colour)
```

#### Exploration hypercube

\hfill

We subsequently define a region near the MLE (hypercube) from which we draw
several hundreds of samples. The specific number of samples varies according to
the complexity of exploring each parameter's space. For $\zeta$, we draw 1,000 
samples.

\hfill


```{r}
loglik_df |>  
  filter(loglik > max(loglik)- 20, loglik.se < 2) |> 
  sapply(range) -> box

set.seed(848406297)

profile_design(
  par_beta  = seq(0.5, 2,length = 50),
  lower     = box[1, c("I_0", "par_rho", "par_inv_phi")],
  upper     = box[2, c("I_0", "par_rho", "par_inv_phi")],
  nprof     = 20, type = "sobol"
) -> guesses
```

```{r, message = FALSE, fig.height = 3, fig.width = 3.5, fig.align = 'center'}
plot_guesses(guesses, 0.25, GBM_colour)
```

#### Likelihood maximisation

\hfill

```{r}
beta_ptb <- rw_sd(I_0 = ivp(0.02), par_inv_phi = 0.02, par_rho = 0.02)

fn_ifp     <- file.path(folder, "ifp_par_beta.rds" )

ifp_obj <- iter_filt_profile(mf1 = mf1, 
                             guesses = guesses,
                             fixed_params = fixed_params,
                             perturbations = beta_ptb,
                             filename = fn_ifp,
                             seed = 23485074, n_cores = detectCores() - 1)
```

```{r, message = FALSE, fig.height = 3.5}
profile_mif_results <- extract_mif_results(ifp_obj)

profile_mif_results |>  
  traces() |> 
  melt() -> traces_zeta_df 

mif_traces(traces_zeta_df, names(par_obj$unknown), GBM_colour)
```